Modeling and Simulation of Load Balancing Strategies for Computing in High Energy Physics

### 2018-10-09 paper/paper.tex
Fix typo
Rework validation section

@@

\section{Validation}
\label{validation}
To validate the model and simulation we simulated CMS jobs at GridKa in April 2018.
The job performance data was randomly separated into two equal sized calibration and test datasets. For the validation we considered the following metrics: Overall throughput of jobs, shares of the job types and the CPU efficiency of the site. These metrics describe how well the simulation matches the measured performance data.

The simulated throughput of the test dataset matches the measured one accurately with only 1\% difference.
\Cref{shares} depicts the share of the simulated and measured job types that also shows  a high accuracy.

\begin{figure*}[h]
	\centering
	\includegraphics[scale = 0.6]{images/shares}
	\caption[]{Measured and Simulated Shares of Job Types}
	\label{shares}
\end{figure*}

The simulated CPU efficiency is lower than the measured one, as shown in \Cref{utilization}. The reason for this are likely missing jobs in the job data. We observed that a significant amount of data is missing, the cause for this is not known yet, however this should be investigated. Because of that the calibrated throughput is too low, resulting in a lower utilization. Another reason is the missing utilization created by pilot jobs that are not explicitly modeled yet.

\begin{figure*}[h]
	\vspace{-0.5ex}
	\centering
	\begin{tikzpicture}[scale = 0.8]
	\begin{axis}[
	xlabel={Time / s},
	ylabel={Total CPU Time / Walltime},
	ymin=0, ymax=1,
	xmin=0, xmax=604800,
	legend cell align={left},
	]
	\addplot[color=red] coordinates {
		(0, 0.61)			
		(604800, 0.61)
	};
	\addplot[color=blue] coordinates {	
		(14000,0.420457263335)
		(15000,0.455205081375)
		(16000,0.450933749429)
		(17000,0.46853252847)
		(18000,0.474029527868)
		(19000,0.482875361196)
		(20000,0.472149795914)
		(21000,0.448813851682)
		(22000,0.430706201872)
		(23000,0.401599999586)
		(24000,0.376755783861)
		(25000,0.38749824392)
		(26000,0.422990841095)
		(27000,0.425760711131)
		(28000,0.43904605249)
		(29000,0.437695380408)
		(30000,0.433314619731)
		(31000,0.448449713155)
		(32000,0.441020436611)
		(33000,0.444429125129)
		(34000,0.461008578928)
		(35000,0.470284257719)
		(36000,0.447847515795)
		(37000,0.441187264182)
		(38000,0.433373025865)
		(39000,0.426778346)
		(40000,0.440334830361)
		(41000,0.452643935542)
		(42000,0.452142136739)
		(43000,0.452471374183)
		(44000,0.445302566248)
		(45000,0.437814297445)
		(46000,0.445663557997)
		(47000,0.451677101561)
		(48000,0.442604461856)
		(49000,0.453882226309)
		(50000,0.444581524102)
		(51000,0.443425493542)
		(52000,0.424588550506)
		(53000,0.424918434251)
		(54000,0.438754707788)
		(55000,0.431586617405)
		(56000,0.432664270787)
		(57000,0.431460959005)
		(58000,0.43776201488)
		(59000,0.452017801875)
		(60000,0.445027944703)
		(61000,0.447754439714)
		(62000,0.441042254855)
		(63000,0.448994039435)
		(64000,0.447949217961)
		(65000,0.440335544034)
		(66000,0.432872781216)
		(67000,0.444765956656)
		(68000,0.445503771932)
		(69000,0.435313479888)
		(70000,0.440723673441)
		(71000,0.453581869677)
		(72000,0.455011550524)
		(73000,0.445251024271)
		(74000,0.461605049304)
		(75000,0.470460077846)
		(76000,0.467995526805)
		(77000,0.495748712572)
		(78000,0.49681048222)
		(79000,0.495635118813)
		(80000,0.476519088357)
		(81000,0.469046301151)
		(82000,0.459260808853)
		(83000,0.443186601975)
		(84000,0.453927784554)
		(85000,0.460996975902)
		(86000,0.439846472086)
		(87000,0.42853879571)
		(88000,0.425940455445)
		(89000,0.418188095172)
		(90000,0.441124921832)
		(91000,0.430042869457)
		(92000,0.454088618748)
		(93000,0.447812967768)
		(94000,0.472828979798)
		(95000,0.503483458242)
		(96000,0.49269529217)
		(97000,0.485731894357)
		(98000,0.500299990804)
		(99000,0.489833797954)
		(100000,0.486536917579)
		(101000,0.470900103725)
		(102000,0.473487499528)
		(103000,0.472852199897)
		(104000,0.465907927377)
		(105000,0.46843843884)
		(106000,0.469393804824)
		(107000,0.448337382107)
		(108000,0.453458598472)
		(109000,0.477909863511)
		(110000,0.484074360628)
		(111000,0.471446789672)
		(112000,0.483311476858)
		(113000,0.492794346306)
		(114000,0.494774125893)
		(115000,0.478760296453)
		(116000,0.463289501118)
		(117000,0.470560118374)
		(118000,0.46261989666)
		(119000,0.448321324595)
		(120000,0.4391456597)
		(121000,0.44616088099)
		(122000,0.443067693493)
		(123000,0.443765615997)
		(124000,0.45489288544)
		(125000,0.457761907159)
		(126000,0.459890981374)
		(127000,0.457265336864)
		(128000,0.450661574709)
		(129000,0.456319671876)
		(130000,0.455222292845)
		(131000,0.444168136077)
		(132000,0.455292020739)
		(133000,0.459739659432)
		(134000,0.474266033943)
		(135000,0.476719078979)
		(136000,0.454265351086)
		(137000,0.432403127216)
		(138000,0.426941224365)
		(139000,0.435303571358)
		(140000,0.445594945267)
		(141000,0.457945319262)
		(142000,0.467041094326)
		(143000,0.471275880951)
		(144000,0.448803750886)
		(145000,0.46057752759)
		(146000,0.450222072137)
		(147000,0.447748079045)
		(148000,0.463476880757)
		(149000,0.457547341501)
		(150000,0.464860371781)
		(151000,0.458939834833)
		(152000,0.449435321435)
		(153000,0.445809287248)
		(154000,0.452098810214)
		(155000,0.436323981134)
		(156000,0.435163726955)
		(157000,0.430055068196)
		(158000,0.432586012365)
		(159000,0.437297849322)
		(160000,0.424896762305)
		(161000,0.433031154001)
		(162000,0.42905370657)
		(163000,0.412624449914)
		(164000,0.412491612196)
		(165000,0.399704545922)
		(166000,0.407748045902)
		(167000,0.409904586648)
		(168000,0.412048703271)
		(169000,0.422106417772)
		(170000,0.431756334979)
		(171000,0.432552532485)
		(172000,0.461650235913)
		(173000,0.450819280269)
		(174000,0.455592609364)
		(175000,0.461503577581)
		(176000,0.463421686809)
		(177000,0.456182986217)
		(178000,0.465859886059)
		(179000,0.469016662103)
		(180000,0.457242602407)
		(181000,0.438524914279)
		(182000,0.419623185788)
		(183000,0.434867384803)
		(184000,0.436354031073)
		(185000,0.431955984805)
		(186000,0.426743598615)
		(187000,0.433639371186)
		(188000,0.447367993742)
		(189000,0.441424080277)
		(190000,0.460353123518)
		(191000,0.460813912595)
		(192000,0.455819448956)
		(193000,0.467462336883)
		(194000,0.475473651551)
		(195000,0.464487515787)
		(196000,0.45335938384)
		(197000,0.435968082782)
		(198000,0.459905581295)
		(199000,0.443261087286)
		(200000,0.438900909062)
		(201000,0.470264321796)
		(202000,0.451214857632)
		(203000,0.468171752558)
		(204000,0.459209586602)
		(205000,0.459254298111)
		(206000,0.47101680367)
		(207000,0.463033170114)
		(208000,0.4712737853)
		(209000,0.465450336139)
		(210000,0.455801602051)
		(211000,0.461908888488)
		(212000,0.458205085034)
		(213000,0.461468054298)
		(214000,0.453173854006)
		(215000,0.454871547866)
		(216000,0.444018420001)
		(217000,0.460956695722)
		(218000,0.465661865182)
		(219000,0.465518305154)
		(220000,0.4711530613)
		(221000,0.459010429226)
		(222000,0.464906045582)
		(223000,0.453027764206)
		(224000,0.455831376049)
		(225000,0.469807813442)
		(226000,0.456897450438)
		(227000,0.446722893031)
		(228000,0.444441981548)
		(229000,0.431491369988)
		(230000,0.448834255688)
		(231000,0.445076241297)
		(232000,0.435089055194)
		(233000,0.437592817834)
		(234000,0.437875351009)
		(235000,0.434588735814)
		(236000,0.44507195818)
		(237000,0.439254948215)
		(238000,0.444706733582)
		(239000,0.447474042265)
		(240000,0.446329940764)
		(241000,0.455749636545)
		(242000,0.419361606614)
		(243000,0.445397279873)
		(244000,0.452348733449)
		(245000,0.434373580372)
		(246000,0.43809435824)
		(247000,0.457201403905)
		(248000,0.458802551256)
		(249000,0.440383453863)
		(250000,0.445017392448)
		(251000,0.450084408166)
		(252000,0.441726249534)
		(253000,0.43194266247)
		(254000,0.445568321745)
		(255000,0.436278081802)
		(256000,0.439570299584)
		(257000,0.445136501732)
		(258000,0.45903255575)
		(259000,0.450554481769)
		(260000,0.446171160379)
		(261000,0.458690240049)
		(262000,0.470220369345)
		(263000,0.480684091886)
		(264000,0.456854673542)
		(265000,0.450233371349)
		(266000,0.457899571788)
		(267000,0.453334820977)
		(268000,0.460295743577)
		(269000,0.457349902193)
		(270000,0.462039377481)
		(271000,0.481382348042)
		(272000,0.491382702182)
		(273000,0.484220311688)
		(274000,0.491652039182)
		(275000,0.486379984979)
		(276000,0.494978853484)
		(277000,0.490824451822)
		(278000,0.461792687666)
		(279000,0.469822396762)
		(280000,0.48576001703)
		(281000,0.481488945743)
		(282000,0.480743747455)
		(283000,0.486090522042)
		(284000,0.487356854396)
		(285000,0.465745257924)
		(286000,0.472330225888)
		(287000,0.492995112907)
		(288000,0.489510571313)
		(289000,0.493418264292)
		(290000,0.489755865611)
		(291000,0.489256145131)
		(292000,0.483996042192)
		(293000,0.471077668813)
		(294000,0.471398859559)
		(295000,0.461127782716)
		(296000,0.458754727805)
		(297000,0.451312728894)
		(298000,0.457666739683)
		(299000,0.463072024695)
		(300000,0.452352488238)
		(301000,0.432056039316)
		(302000,0.442534543126)
		(303000,0.464020942755)
		(304000,0.465788968846)
		(305000,0.475194751884)
		(306000,0.460429089784)
		(307000,0.440203075328)
		(308000,0.440392507843)
		(309000,0.469031285375)
		(310000,0.457737564909)
		(311000,0.450884070739)
		(312000,0.467259163074)
		(313000,0.467018061082)
		(314000,0.466325864949)
		(315000,0.4515369024)
		(316000,0.451628960618)
		(317000,0.449715736896)
		(318000,0.453599972995)
		(319000,0.46598650055)
		(320000,0.458071641171)
		(321000,0.459795875349)
		(322000,0.463776572273)
		(323000,0.441893234874)
		(324000,0.450990407649)
		(325000,0.469838695448)
		(326000,0.454971976955)
		(327000,0.458423340158)
		(328000,0.455021879641)
		(329000,0.474962772272)
		(330000,0.468921434665)
		(331000,0.463586286578)
		(332000,0.472336338416)
		(333000,0.480203877337)
		(334000,0.499976146642)
		(335000,0.499528976694)
		(336000,0.486582556042)
		(337000,0.486946662392)
		(338000,0.472474569197)
		(339000,0.47914245784)
		(340000,0.475286986461)
		(341000,0.458659120511)
		(342000,0.463977117388)
		(343000,0.48570346922)
		(344000,0.47347942112)
		(345000,0.450392499597)
		(346000,0.445409073982)
		(347000,0.454343590968)
		(348000,0.446283787024)
		(349000,0.441991195931)
		(350000,0.457080478112)
		(351000,0.475143781352)
		(352000,0.453825073455)
		(353000,0.466696828065)
		(354000,0.461398654993)
		(355000,0.476432737168)
		(356000,0.484493658032)
		(357000,0.478381386037)
		(358000,0.48791767455)
		(359000,0.471135835878)
		(360000,0.4510006391)
		(361000,0.459254358243)
		(362000,0.45940296355)
		(363000,0.455276346902)
		(364000,0.460372976975)
		(365000,0.46432780643)
		(366000,0.468703252372)
		(367000,0.460690627792)
		(368000,0.456596843643)
		(369000,0.453188747838)
		(370000,0.444026572421)
		(371000,0.469540885626)
		(372000,0.477111174668)
		(373000,0.453778831255)
		(374000,0.451432024952)
		(375000,0.467745860758)
		(376000,0.488699820134)
		(377000,0.47828379946)
		(378000,0.470780021427)
		(379000,0.460858540737)
		(380000,0.468116574189)
		(381000,0.476739105023)
		(382000,0.465550165705)
		(383000,0.466704565487)
		(384000,0.460951873868)
		(385000,0.474085333555)
		(386000,0.483774319139)
		(387000,0.450257163363)
		(388000,0.432558156082)
		(389000,0.441761962984)
		(390000,0.471357238783)
		(391000,0.476402662876)
		(392000,0.478773054016)
		(393000,0.479343982261)
		(394000,0.476955005539)
		(395000,0.464686525062)
		(396000,0.465867162002)
		(397000,0.473713210956)
		(398000,0.455977533182)
		(399000,0.453556194091)
		(400000,0.468032168961)
		(401000,0.473372878696)
		(402000,0.460103363611)
		(403000,0.450437880682)
		(404000,0.444189180728)
		(405000,0.440631553941)
		(406000,0.440814465656)
		(407000,0.439132208745)
		(408000,0.443680589194)
		(409000,0.439364305134)
		(410000,0.445754620988)
		(411000,0.459675216489)
		(412000,0.434411774646)
		(413000,0.460859998054)
		(414000,0.470068659819)
		(415000,0.46398628613)
		(416000,0.46092602178)
		(417000,0.470952398074)
		(418000,0.464918768054)
		(419000,0.457250642935)
		(420000,0.452696999147)
		(421000,0.443597876602)
		(422000,0.442688896448)
		(423000,0.446975959689)
		(424000,0.459962628411)
		(425000,0.458639133149)
		(426000,0.45869214177)
		(427000,0.459339918847)
		(428000,0.469954533509)
		(429000,0.466801073197)
		(430000,0.479585876798)
		(431000,0.473889297072)
		(432000,0.468945113898)
		(433000,0.46819782837)
		(434000,0.460713604747)
		(435000,0.467085533437)
		(436000,0.464877942532)
		(437000,0.441101434716)
		(438000,0.435732064901)
		(439000,0.448217580287)
		(440000,0.447917450562)
		(441000,0.443407621096)
		(442000,0.443456810485)
		(443000,0.438987312155)
		(444000,0.467286398042)
		(445000,0.469189496526)
		(446000,0.467728960254)
		(447000,0.469175237549)
		(448000,0.469335300568)
		(449000,0.463820736067)
		(450000,0.458428492443)
		(451000,0.448441949348)
		(452000,0.453900123964)
		(453000,0.459953017199)
		(454000,0.448381040252)
		(455000,0.465700695725)
		(456000,0.451531261356)
		(457000,0.456563899123)
		(458000,0.45720539038)
		(459000,0.451801077935)
		(460000,0.464657248779)
		(461000,0.459373104135)
		(462000,0.469958682024)
		(463000,0.467903777714)
		(464000,0.445428785915)
		(465000,0.442499436959)
		(466000,0.447134373326)
		(467000,0.438282409091)
		(468000,0.43839540036)
		(469000,0.419945925767)
		(470000,0.418452547982)
		(471000,0.438307200322)
		(472000,0.45350446074)
		(473000,0.468786686762)
		(474000,0.456254084131)
		(475000,0.453297591133)
		(476000,0.469078411526)
		(477000,0.457354866601)
		(478000,0.476430954364)
		(479000,0.475029005373)
		(480000,0.479985350112)
		(481000,0.475381159607)
		(482000,0.461543267652)
		(483000,0.451592900155)
		(484000,0.45260612709)
		(485000,0.458008617252)
		(486000,0.444827616864)
		(487000,0.447400533469)
		(488000,0.452399703406)
		(489000,0.451208289224)
		(490000,0.469255926366)
		(491000,0.469056554934)
		(492000,0.448427595423)
		(493000,0.460791545132)
		(494000,0.465807144119)
		(495000,0.459466948041)
		(496000,0.457398936667)
		(497000,0.454270854364)
		(498000,0.44030861198)
		(499000,0.441478571167)
		(500000,0.448416540804)
		(501000,0.447708055398)
		(502000,0.44394661343)
		(503000,0.437604010563)
		(504000,0.43279917301)
		(505000,0.448202463548)
		(506000,0.450335081656)
		(507000,0.456367109826)
		(508000,0.476797822515)
		(509000,0.464004036098)
		(510000,0.447462462599)
		(511000,0.460128351384)
		(512000,0.456610531662)
		(513000,0.461027700674)
		(514000,0.452647871516)
		(515000,0.472383445652)
		(516000,0.480048107691)
		(517000,0.449495011852)
		(518000,0.424921363583)
		(519000,0.430957570675)
		(520000,0.42946792577)
		(521000,0.435679220854)
		(522000,0.454543429655)
		(523000,0.449873483822)
		(524000,0.441368528175)
		(525000,0.434726970329)
		(526000,0.446558845708)
		(527000,0.436523566317)
		(528000,0.436567752145)
		(529000,0.440939357596)
		(530000,0.442182493949)
		(531000,0.446092107838)
		(532000,0.451187916749)
		(533000,0.439896893747)
		(534000,0.449011137767)
		(535000,0.444120664286)
		(536000,0.426232253223)
		(537000,0.438918286378)
		(538000,0.44181502269)
		(539000,0.432995321972)
		(540000,0.434879809332)
		(541000,0.427145288337)
		(542000,0.427872289185)
		(543000,0.417701942996)
		(544000,0.419515873503)
		(545000,0.439790579048)
		(546000,0.43762292406)
		(547000,0.443042538052)
		(548000,0.455727191114)
		(549000,0.447614026634)
		(550000,0.456624751487)
		(551000,0.452793939028)
		(552000,0.440726166681)
		(553000,0.449038613212)
		(554000,0.468536381832)
		(555000,0.444002288289)
		(556000,0.452092079245)
		(557000,0.456883014276)
		(558000,0.453112451125)
		(559000,0.441037791264)
		(560000,0.442347996342)
		(561000,0.436264841204)
		(562000,0.438959132682)
		(563000,0.4451724287)
		(564000,0.458215756554)
		(565000,0.448690251696)
		(566000,0.453197454907)
		(567000,0.453794387302)
		(568000,0.444846550094)
		(569000,0.449982902679)
		(570000,0.447267180536)
		(571000,0.445828406211)
		(572000,0.446909040851)
		(573000,0.453980157655)
		(574000,0.457292498947)
		(575000,0.448977443253)
		(576000,0.446078471261)
		(577000,0.434934699022)
		(578000,0.45949539644)
		(579000,0.477069726029)
		(580000,0.478983629734)
		(581000,0.459161976806)
		(582000,0.470964357654)
		(583000,0.466734817389)
		(584000,0.451817556376)
		(585000,0.438500345769)
		(586000,0.435540269671)
		(587000,0.436722886274)
		(588000,0.452696266979)
		(589000,0.444922652517)
		(590000,0.414546052368)
		(591000,0.40666087591)
		(592000,0.431870620823)
		(593000,0.456065081956)
		(594000,0.45959389509)
		(595000,0.445709712809)
		(596000,0.455310403564)
		(597000,0.474126321375)
		(598000,0.466945015295)
		(599000,0.442916390259)
		(600000,0.463327317157)
		(601000,0.463538228748)
		(602000,0.484414072165)
		(603000,0.470969958228)
		(604000,0.357975119833)

	};
	\addlegendentry{Measured (average)};
	\addlegendentry{Simulated};
	\end{axis}
	\end{tikzpicture}
	\caption[]{Measured and Simulated CPU Efficiency}
	\label{utilization}
\end{figure*}

Simulating one week of CMS jobs at GridKa takes about 24 minutes on a laptop. Thus, it would feasible to simulate several Tiers at once in a sufficient time, however simulating larger parts of the grid is future work.

@@

We successfully simulated the CMS computing jobs at GridKa. To enable easy usage of the simulation we implemented an automated model parameter extraction and model creation. Furthermore, we extended Palladio to be able to simulate load balancing strategies. Our first results are promising and suggest that this approach can be used to simulate load balancing strategies accurately. Regarding scalability we are confident that several computing centers can simulated in a sufficient time, however this is part of future work.

>>>>

% Use package for subfigures
\usepackage{subcaption}

@@

\section{Validation}
\label{validation}

\begin{figure}
    \centering
    \begin{subfigure}{0.48\linewidth}
        \includegraphics[width=\linewidth]{images/validation/throughputs}
    \end{subfigure}
    \begin{subfigure}{0.48\linewidth}
        \includegraphics[width=\linewidth]{images/validation/utilizations}
    \end{subfigure}
    \caption{Measured and simulated throughputs and CPU efficiency for test data from May 2018}
    \label{fig:validation}
\end{figure}

To validate the model and simulation we simulated CMS jobs at GridKa in May 2018.
The job performance data was split into two equal sized calibration and test datasets by randomly sampling reports from the full data set.
For the validation we considered the following metrics: Overall throughput of jobs, shares of the job types, and the overall CPU efficiency of the site.
These metrics describe how well the simulation matches the measured performance data.
Comparisons between  simulation results and measurements from the system itself for them are shown in figure \ref{fig:validation}.
The simulated throughput matches the reference from the with a difference of about 5.6\%.
The share of the simulated and measured job types also match accurately.

The CPU utilization in the simulation and a comparison with the measured CPU efficiency is also shown in figure \ref{fig:validation}.
Here, the ramp-up of the simulation is also visible.
The overall average utilization in the simulation deviates from the reference by about 3\%, showing that even the current simplified model is capable of accurately reproducing the metrics introduced above.
Runtimes of the simulated jobs were also measured, the average runtime of simulated jobs deviates by approximately 8\% from the measured runtimes in the test data set.

Simulating one week of CMS jobs at GridKa takes about 24 minutes on a laptop.
Thus, it would be feasible to simulate several sites at once in a reasonable time, however simulating larger parts of the grid is future work.

@@

We successfully simulated the CMS computing jobs at GridKa. To enable easy usage of the simulation we implemented an automated model parameter extraction and model creation. Furthermore, we extended Palladio to be able to simulate load balancing strategies. Our first results are promising and suggest that this approach can be used to simulate load balancing strategies accurately. Regarding scalability we are confident that several computing centers can be simulated in a sufficient time, however this is part of future work.
